---
title: Raw Maternity Leave Analysis
author: Katherine Taylor
output: html_document
---

```{r setup, include = FALSE}
# options
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
options(gt.quiet = TRUE)

# seed
set.seed(91006)
# libraries
library(tidyverse)
library(janitor)
library(here)
library(hrbrthemes)
library(modelsummary)
library(gt)
#hrbrthemes::import_roboto_condensed()

# global settings
gof_stuff <- tribble(
  ~raw, ~clean, ~fmt,
  "nobs", "N", 0,
  "r.squared", "R²", 3
)
```

# Data Cleaning
```{r read_in_data}
mat_health_raw <-
  read_csv(here("data/raw_data/maternal_health_outcomes.csv")) |> clean_names()
mat_leave_raw <-
  read_csv(here("data/raw_data/maternity_leave_raw_data.csv")) |> clean_names()
```

Checks that I would do 
- Is the data for every country in weeks?
- What is the most recent year in the data, is it all the same?
- Are the columns sex and sex_2 and time and time_2 the same? If not, why are they different?
- Is there the same number of countries before and after the filter?

```{r data_checks}
length(unique(mat_leave_raw$country))
```

```{r clean_leave_data}
mat_leave_clean <- mat_leave_raw |>
  filter(indicator == "Length of maternity leave" &
           sex == "WOMEN") |>
  group_by(country) |>
  filter(cou != "OAVG") |>
  ungroup()
```

```{r post_cleaning_checks}
# should be raw - 1 since I'm removing the average point
length(unique(mat_leave_clean$country))

# look for outliers in the maternity leave value column
ggplot(mat_leave_clean, aes(x = value)) +
  geom_density() +
  theme_ipsum()
```

```{r get_countries_list}
countries <- mat_leave_clean |> select(country, time) |> unique()
```

```{r clean_outcomes_data}
mat_health_clean <- mat_health_raw |>
  filter(time_period_time_period %in% countries$time) |>
  mutate(
    country_cleaned = str_trim(str_remove_all(ref_area_geographic_area, "^[^*]*:")),
    time = parse_number(time_period_time_period)
  ) |>
  filter(country_cleaned %in% countries$country)

final_countries <-
  data.frame("country" = unique(mat_health_clean$country_cleaned),
             "num" = 1:34)
```

```{r final_df}
mat_df <- mat_leave_clean |>
  left_join(mat_health_clean,
            by = c("country" = "country_cleaned", "time")) |>
  filter(
    indicator_indicator == "MNCH_MMR: Maternal mortality ratio (number of maternal deaths per 100,000 live births)" &
      sex_sex == "F: Female"
  ) |>
  select(
    cou,
    country,
    time,
    value,
    obs_value_observation_value,
    unit_measure_unit_of_measure
  ) |>
  rename(
    "year" = "time",
    "weeks_mat_leave" = "value",
    "mat_mort_rate" = "obs_value_observation_value",
    "unit" = "unit_measure_unit_of_measure"
  ) |>
  mutate(mat_mort_rate = parse_number(mat_mort_rate),
         mat_mort_rate = mat_mort_rate / 100000)
```

```{r save_cleaned_data}
write_csv(mat_df, here("data/clean_data/maternity_data_cleaned.csv"))
save(mat_df, file = here("data/clean_data/maternity_data_cleaned.RData"))
```

# Data Exploration
```{r}
skimr::skim(mat_df)
```

# Data Visualization
```{r}
mat_df |> 
  ggplot(aes(x = weeks_mat_leave, y = mat_mort_rate, color = country)) +
  geom_point() +
  facet_grid(~year)
```

```{r}
mat_df |> 
  ggplot(aes(x = year, y = mat_mort_rate, color = country)) +
  geom_line()
```

# Linear Model
```{r}
model_ols_1 <- lm(mat_mort_rate ~ weeks_mat_leave, data = mat_df)
model_ols_2 <- lm(mat_mort_rate ~ year, data = mat_df)
model_ols_3 <-
  lm(mat_mort_rate ~ weeks_mat_leave + year, data = mat_df)
```

```{r}
modelsummary(list(model_ols_1, model_ols_2, model_ols_3), gof_map = gof_stuff, output = "gt")
```

```{r}
par(mfrow = c(2,2))
plot(model_ols_1)
plot(model_ols_2)
plot(model_ols_3)
par(mfrow = c(1,1))
```
There are clearly some time related trends that need to be taken care of at some point. Also, the data is nowhere close to fitting a normal distribution. 
#TODO reread time series book on lagged regression

# GLM Model
```{r}
model_frac_logit1 <- glm(mat_mort_rate ~ weeks_mat_leave, data = mat_df,
                         family = quasibinomial())
model_frac_logit2 <- glm(mat_mort_rate ~ year, data = mat_df,
                         family = quasibinomial())
model_frac_logit3 <- glm(mat_mort_rate ~ weeks_mat_leave + year, data = mat_df,
                         family = quasibinomial())
```

```{r}
modelsummary(list(model_frac_logit1, model_frac_logit2, model_frac_logit3),gof_map = gof_stuff, output = "gt")
```

```{r}
summary(model_frac_logit3)
plot(model_frac_logit3)
```

# Test adding country to the model
```{r}
model_frac_logitc <- glm(mat_mort_rate ~ weeks_mat_leave + year + country, data = mat_df,
                         family = quasibinomial())
summary(model_frac_logitc)
plot(model_frac_logitc)
```

